Object Oriented Shader Composition Using CLOS

### 2018-04-04 els2018-glsl-oop/paper.tex
Add some stuff about applications.

\begin{figure}[h]
  \begin{center}
    \begin{minipage}{.2\textwidth}
      \includegraphics[width=0.9\textwidth]{render-regular.png}
    \end{minipage}
    \begin{minipage}{.2\textwidth}
      \includegraphics[width=0.9\textwidth]{render-warped.png}
    \end{minipage}
  \end{center}
  \caption{Left: rendering of a sphere using \autoref{lst:vertex}. Right: the same with \autoref{lst:warp} added.}
  \label{fig:regular}
\end{figure}

Using this technique, a wide variety of shaders can be combined, with a minimal amount of awareness of other shaders being necessary. Shortcomings of this technique are elaborated in \autoref{section:5}.

@@

\section{Conclusion}\label{section:5}
Source code analysis allows us to easily re-use shader code written independently of our Lisp program while retaining the ability to merge the effects of multiple shaders together. Integrating this analysis into the object system combines the GPU and CPU logic in a single class, allowing the inheritance and re-use of both. \\

@@

\section{Further Work}\label{section:6}
The current merging strategy employed is rather simplistic. For instance, no attempt at recognising identical function definitions is made. The system could also be extended with a variety of static analysis algorithms to optimise the code ahead of time, or to provide errors and warnings about potential user mistakes. \\

@@

\section{Acknowledgements}\label{section:7}
I would like to thank Robert Strandh and Philipp Marek for feedback on the paper.

\section{Implementation}\label{section:8}
An implementation of the proposed system can be found at \href{https://github.com/Shirakumo/trial/blob/e37e0dc73085c401e43da58d5098a9cf02167f8f/shader-entity.lisp}{https://github.com/Shirakumo/trial/blob/\\e37e0dc73085c401e43da58d5098a9cf02167f8f/shader-entity.lisp} for the CLOS part, and at \url{https://github.com/Shirakumo/glsl-toolkit} for the merger part. \\

>>>>

\begin{figure}[h]
  \begin{center}
    \begin{minipage}{.2\textwidth}
      \includegraphics[width=1.0\textwidth]{render-regular.png}
    \end{minipage}
    \begin{minipage}{.2\textwidth}
      \includegraphics[width=1.0\textwidth]{render-warped.png}
    \end{minipage}
  \end{center}
  \caption{Left: rendering of a sphere using \autoref{lst:vertex}. Right: the same with \autoref{lst:warp} added.}
  \label{fig:regular}
\end{figure}

Using this technique, a wide variety of shaders can be combined, with a minimal amount of awareness of other shaders being necessary. Shortcomings of this technique are elaborated in \autoref{section:6}.

@@

\section{Applications}\label{section:5}
We currently use this technique for two primary applications: the encapsulation of common display functionality into mixins, and the implementation of effects passes. The former allows us to separate various concerns such as the read-out of texture data, the transformation of vertex data, and the fragment rendering behaviour into small, self-contained classes, that can then be combined by the user to produce the result they desire. \\

The latter allows us to write arbitrary rendering effects as self-contained render pass objects. To elaborate on this application, imagine a visual effect that requires you to manipulate some, but not all of the aspects of how an object is rendered. A simple example for such an effect is the ``god rays'' that you can see in \autoref{fig:godrays}. \\

In order to achieve this effect, all opaque objects in the scene are rendered once normally, and once in black. Then the black pass is blurred and added to the normal render. Neither the normal render pass nor the combination step require any insight from the pass into how the objects are drawn. However, for the black rendering pass, the way the objects are drawn has to be modified in such a way that all fragments are output as black, but everything else, such as vertex transformations, must remain the same. \\

With shader combination we can achieve this quite easily: we write a fragment shader that simply outputs the colour black, and then combine this shader with whatever the objects' shaders are. In this way we are virtually inserting a superclass for each object in the scene depending on which render pass is currently active. Render passes themselves are implemented with the same class shader combination as regular objects, so they too can profit from encapsulating and sharing shader functionality.

\begin{figure}[h]
  \begin{center}
    \begin{minipage}{.2\textwidth}
      \includegraphics[width=1.0\textwidth]{pass-render.png}
    \end{minipage}
    \begin{minipage}{.2\textwidth}
      \includegraphics[width=1.0\textwidth]{pass-black.png}
    \end{minipage}
    \vskip0.1cm
    \begin{minipage}{.4055\textwidth}
      \includegraphics[width=1.0\textwidth]{pass-combined.png}
    \end{minipage}
  \end{center}
  \caption{Left: phong rendered teapot. Right: black rendered teapot. Bottom: combined god-rays effect.}
  \label{fig:godrays}
\end{figure}

\section{Conclusion}\label{section:6}
Source code analysis allows us to easily re-use shader code written independently of our Lisp program while retaining the ability to merge the effects of multiple shaders together. Integrating this analysis into the object system combines the GPU and CPU logic in a single class, allowing the inheritance and re-use of both. \\

@@

\section{Further Work}\label{section:7}
The current merging strategy employed is rather simplistic. For instance, no attempt at recognising identical function definitions is made. The system could also be extended with a variety of static analysis algorithms to optimise the code ahead of time, or to provide errors and warnings about potential user mistakes. \\

@@

\section{Acknowledgements}\label{section:8}
I would like to thank Robert Strandh and Philipp Marek for feedback on the paper.

\section{Implementation}\label{section:9}
An implementation of the proposed system can be found at \href{https://github.com/Shirakumo/trial/blob/e37e0dc73085c401e43da58d5098a9cf02167f8f/shader-entity.lisp}{https://github.com/Shirakumo/trial/blob/\\e37e0dc73085c401e43da58d5098a9cf02167f8f/shader-entity.lisp} for the CLOS part, and at \url{https://github.com/Shirakumo/glsl-toolkit} for the merger part. \\
